A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2024-01-16 tidyindex.qmd
modify patricia's affiliation

---
title: "A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data"
format:
  tandf-pdf:
    keep-tex: true  
    linestretch: 2
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-3813-7155
  - name: Ursula Laa
    affiliation: 
      - name: University of Natural Resources and Life Sciences
        department: Institute of Statistics
        address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
        city: Vienna
        country: Austria
    orcid: 0000-0002-0249-6439
  - name: Nicolas Langrené
    affiliation: 
      - name: BNU-HKBU United International College
        department: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science
        address: 2000 Jintong Road, Tangjiawan, Zhuhai
        city: Zhuhai, Guangdong
        country: China
    orcid: 0000-0001-7601-4618
  - name: Patricia Menéndez
    affiliation:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0003-0701-6315
abstract: |
  Indexes are useful for summarizing multivariate information into single metrics for monitoring, communicating, and decision-making. While most work has focused on defining new indexes for specific purposes, more attention needs to be directed towards making it possible to understand index behavior in different data conditions, and to determine how their structure affects their values and variation in values. Here we discuss a modular data pipeline recommendation to assemble indexes. It is universally applicable to index computation and allows investigation of index behavior as part of the development procedure. One can compute indexes with different parameter choices, adjust steps in the index definition by adding, removing, and swapping them to experiment with various index designs, calculate uncertainty measures, and assess indexes' robustness. The paper presents three examples to illustrate the pipeline framework usage: comparison of two different indexes designed to monitor the spatio-temporal distribution of drought in Queensland, Australia; the effect of dimension reduction choices on the Global Gender Gap Index (GGGI) on countries' ranking; and how to calculate bootstrap confidence intervals for the Standardized Precipitation Index (SPI). The methods are supported by a new R package, called `tidyindex`.
keywords:
  - indexes
  - data pipeline
  - software design
  - uncertainty
  - decision-making
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

>>>>

---
title: "A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data"
format:
  tandf-pdf:
    keep-tex: true  
    linestretch: 2
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-3813-7155
  - name: Ursula Laa
    affiliation: 
      - name: University of Natural Resources and Life Sciences
        department: Institute of Statistics
        address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
        city: Vienna
        country: Austria
    orcid: 0000-0002-0249-6439
  - name: Nicolas Langrené
    affiliation: 
      - name: BNU-HKBU United International College
        department: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science
        address: 2000 Jintong Road, Tangjiawan, Zhuhai
        city: Zhuhai, Guangdong
        country: China
    orcid: 0000-0001-7601-4618
  - name: Patricia Menéndez
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
      - name: University of Melbourne
        department: School of Mathematics and Statistics
        address: 813 Swanston Street, Parkville VIC 3052, Australia
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0003-0701-6315
abstract: |
  Indexes are useful for summarizing multivariate information into single metrics for monitoring, communicating, and decision-making. While most work has focused on defining new indexes for specific purposes, more attention needs to be directed towards making it possible to understand index behavior in different data conditions, and to determine how their structure affects their values and variation in values. Here we discuss a modular data pipeline recommendation to assemble indexes. It is universally applicable to index computation and allows investigation of index behavior as part of the development procedure. One can compute indexes with different parameter choices, adjust steps in the index definition by adding, removing, and swapping them to experiment with various index designs, calculate uncertainty measures, and assess indexes' robustness. The paper presents three examples to illustrate the pipeline framework usage: comparison of two different indexes designed to monitor the spatio-temporal distribution of drought in Queensland, Australia; the effect of dimension reduction choices on the Global Gender Gap Index (GGGI) on countries' ranking; and how to calculate bootstrap confidence intervals for the Standardized Precipitation Index (SPI). The methods are supported by a new R package, called `tidyindex`.
keywords:
  - indexes
  - data pipeline
  - software design
  - uncertainty
  - decision-making
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

### 2024-01-16 tidyindex.tex
modify patricia's affiliation

\title{A Tidy Framework and Infrastructure to Systematically Assemble
Spatio-temporal Indexes from Multivariate Data}
\author{H. Sherry
Zhang$\textsuperscript{1}$~\orcidlink{0000-0002-7122-1463}, Dianne
Cook$\textsuperscript{1}$~\orcidlink{0000-0002-3813-7155}, Ursula
Laa$\textsuperscript{2}$~\orcidlink{0000-0002-0249-6439}, Nicolas
Langrené$\textsuperscript{3}$~\orcidlink{0000-0001-7601-4618}, Patricia
Menéndez$\textsuperscript{1}$~\orcidlink{0000-0003-0701-6315}}

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Guangdong Provincial Key
Laboratory of Interdisciplinary Research and Application for Data
Science, BNU-HKBU United International College, Zhuhai, Guangdong, China
\begin{abstract}
Indexes are useful for summarizing multivariate information into single
metrics for monitoring, communicating, and decision-making. While most
work has focused on defining new indexes for specific purposes, more
attention needs to be directed towards making it possible to understand
index behavior in different data conditions, and to determine how their
structure affects their values and variation in values. Here we discuss
a modular data pipeline recommendation to assemble indexes. It is
universally applicable to index computation and allows investigation of
index behavior as part of the development procedure. One can compute
indexes with different parameter choices, adjust steps in the index
definition by adding, removing, and swapping them to experiment with
various index designs, calculate uncertainty measures, and assess
indexes' robustness. The paper presents three examples to illustrate the
pipeline framework usage: comparison of two different indexes designed
to monitor the spatio-temporal distribution of drought in Queensland,
Australia; the effect of dimension reduction choices on the Global
Gender Gap Index (GGGI) on countries' ranking; and how to calculate
bootstrap confidence intervals for the Standardized Precipitation Index
(SPI). The methods are supported by a new R package, called
\texttt{tidyindex}.
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep software design\sep uncertainty\sep 
decision-making
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, enhanced, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

>>>>

\title{A Tidy Framework and Infrastructure to Systematically Assemble
Spatio-temporal Indexes from Multivariate Data}
\author{H. Sherry
Zhang$\textsuperscript{1}$~\orcidlink{0000-0002-7122-1463}, Dianne
Cook$\textsuperscript{1}$~\orcidlink{0000-0002-3813-7155}, Ursula
Laa$\textsuperscript{2}$~\orcidlink{0000-0002-0249-6439}, Nicolas
Langrené$\textsuperscript{3}$~\orcidlink{0000-0001-7601-4618}, Patricia
Menéndez$\textsuperscript{1,4}$~\orcidlink{0000-0003-0701-6315}}

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Guangdong Provincial Key
Laboratory of Interdisciplinary Research and Application for Data
Science, BNU-HKBU United International College, Zhuhai,
Guangdong, China\\ \textsuperscript{4} School of Mathematics and
Statistics, University of Melbourne, Melbourne, Victoria, Australia
\begin{abstract}
Indexes are useful for summarizing multivariate information into single
metrics for monitoring, communicating, and decision-making. While most
work has focused on defining new indexes for specific purposes, more
attention needs to be directed towards making it possible to understand
index behavior in different data conditions, and to determine how their
structure affects their values and variation in values. Here we discuss
a modular data pipeline recommendation to assemble indexes. It is
universally applicable to index computation and allows investigation of
index behavior as part of the development procedure. One can compute
indexes with different parameter choices, adjust steps in the index
definition by adding, removing, and swapping them to experiment with
various index designs, calculate uncertainty measures, and assess
indexes' robustness. The paper presents three examples to illustrate the
pipeline framework usage: comparison of two different indexes designed
to monitor the spatio-temporal distribution of drought in Queensland,
Australia; the effect of dimension reduction choices on the Global
Gender Gap Index (GGGI) on countries' ranking; and how to calculate
bootstrap confidence intervals for the Standardized Precipitation Index
(SPI). The methods are supported by a new R package, called
\texttt{tidyindex}.
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep software design\sep uncertainty\sep 
decision-making
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
